---
title: "Womens History Month ACS/PUMS Data"
author: "suzanne"
date: "3/2022"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Notes for Asian American Pacific Islander month

```{r load_libraries}
library(devtools)
library(sf)
library(dplyr)
library(psrccensus)
library(ggplot2)
library(tidycensus)
library(tidyr)
Sys.getenv("CENSUS_API_KEY")
```
```{r}
acs_5_vars<-load_variables(2020, 'acs5', cache=TRUE)

```


	
B01002D_001 Median Age
	
B02015    specific origins

B03002 Hispanic
	
B08105 Mode to Work
	
B10051 Grandparents
	
B17001 Poverty
	
C15002 Education
	
B16005 English proficiency

RAC1P
RAC2P - inlcudes alone detailed races
RAC3P detailed race in combination
RACASN- alone or in combination

PERNP- earnings

HINCP

HUGCL
HUPAC
MULTG


### Educational Attainment
```{r}




Asian_df <- get_acs_recs(geography = 'county',
             table.names = c('C15002D'),
             years=c(2019),
             acs.type = 'acs5')%>%filter(name=='Region')

PI_df <- get_acs_recs(geography = 'county',
             table.names = c('C15002E'),
             years=c(2019),
             acs.type = 'acs5')%>%filter(name=='Region')

white_df <- get_acs_recs(geography = 'county',
             table.names = c('C15002H'),
             years=c(2019),
             acs.type = 'acs5')%>%filter(name=='Region')


Asian_PI_df<-merge(Asian_df, PI_df, by='label')
Asian_PI_white_df<-merge(Asian_PI_df, white_df, by='label')


Asian_PI_white_df_formatted<-Asian_PI_white_df %>% select(label,estimate.x, moe.x, estimate.y, moe.y, estimate.y, moe.y, estimate, moe) %>% rename(Asian_Educational_Attainment=estimate.x, Asian_Educational_Attainment_MOE=moe.x, Pacific_Islander_Educational_Attainment=estimate.y, Pacific_Islander_Educational_Attainment_MOE=moe.y, White_Educational_Attainment=estimate, White_Educational_Attainment_MOE=moe)





write.table(Asian_PI_white_df_formatted,"clipboard", sep='\t', row.names=FALSE )


Asian_PI_white_df_formatted


```


by county
```{r}




Asian_df <- get_acs_recs(geography = 'county',
             table.names = c('C15002D'),
             years=c(2020),
             acs.type = 'acs5')

PI_df <- get_acs_recs(geography = 'county',
             table.names = c('C15002E'),
             years=c(2020),
             acs.type = 'acs5')

white_df <- get_acs_recs(geography = 'county',
             table.names = c('C15002H'),
             years=c(2020),
             acs.type = 'acs5')


Asian_PI_df<-merge(Asian_df, PI_df, by=c('label', 'name'))
Asian_PI_white_df<-merge(Asian_PI_df, white_df, by=c('label', 'name'))


Asian_PI_white_df_formatted<-Asian_PI_white_df %>% select(name,label,estimate.x, moe.x, estimate.y, moe.y, estimate.y, moe.y, estimate, moe) %>% rename(Asian_Educational_Attainment=estimate.x, Asian_Educational_Attainment_MOE=moe.x, Pacific_Islander_Educational_Attainment=estimate.y, Pacific_Islander_Educational_Attainment_MOE=moe.y, White_Educational_Attainment=estimate, White_Educational_Attainment_MOE=moe)





write.table(Asian_PI_white_df_formatted,"clipboard", sep='\t', row.names=FALSE )


Asian_PI_white_df_formatted



```


### Median Income
B19013D
2019
```{r}




Asian_df <- get_acs_recs(geography = 'county',
             table.names = c('B19013D'),
             years=c(2019),
             acs.type = 'acs5')
PI_df <- get_acs_recs(geography = 'county',
             table.names = c('B19013E'),
             years=c(2019),
             acs.type = 'acs5')

white_df <- get_acs_recs(geography = 'county',
             table.names = c('B19013H'),
             years=c(2019),
             acs.type = 'acs5')


Asian_PI_df<-merge(Asian_df, PI_df, by=c('label', 'name'))
Asian_PI_white_df<-merge(Asian_PI_df, white_df, by=c('label', 'name'))


Asian_PI_white_df_formatted<-Asian_PI_white_df %>% select(label,name, estimate.x, moe.x, estimate.y, moe.y, estimate.y, moe.y, estimate, moe) %>% rename(Asian_Educational_Attainment=estimate.x, Asian_Median_Income_MOE=moe.x, Pacific_Islander_Median_Income=estimate.y, Pacific_Islander_Median_Income_MOE=moe.y, White_Median_Income=estimate, White_Median_Income_MOE=moe)





write.table(Asian_PI_white_df_formatted,"clipboard", sep='\t', row.names=FALSE )


Asian_PI_white_df_formatted
```
## Homeownership

```{r}


homeownership_black<-get_acs_recs(geography = 'county',
             table.names = c('B25003B'),
             years=c(2019),
             acs.type = 'acs1')

homeownership_white<-get_acs_recs(geography = 'county',
             table.names = c('B25003H'),
             years=c(2019),
             acs.type = 'acs1')


homeownership_df_black_white<- merge(homeownership_black, homeownership_white, by=c('label','name'))
write.table(homeownership_df_black_white,"clipboard", sep='\t', row.names=FALSE )

homeownership_df_black_white

```

### Median Age by Black and All

```{r}

Black_age_df<-get_acs_recs(geography = 'county',
             table.names = c('B01002B'),
             years=c(2019),
             acs.type = 'acs5')


white_age_df <- get_acs_recs(geography = 'county',
             table.names = c('B01002H'),
             years=c(2019),
             acs.type = 'acs5')


white_Black_df<-merge(Black_age_df, white_age_df, by=c('name','label'))%>% filter(label=='Estimate!!Median age --!!Total:')%>%
  select(name,estimate.x, estimate.y, )%>% rename('County'='name', 'Black population' ='estimate.x', 'White Population'='estimate.y')

write.table(white_Black_df,"clipboard", sep='\t', row.names=FALSE )
white_Black_df

```



### Poverty Rates

```{r}


poverty_df_Black<- get_acs_recs(geography = 'county',
             table.names = c('B17020B'),
             years=c(2019),
             acs.type = 'acs5')
poverty_df_white <- get_acs_recs(geography = 'county',
             table.names = c('B17020H'),
             years=c(2019),
             acs.type = 'acs5')

poverty_df_white<- poverty_df_white %>% filter(label=='Estimate!!Total:' | label=='Estimate!!Total:!!Income in the past 12 months below poverty level:')%>%select(name, estimate,moe,label)%>%tidyr::pivot_wider(names_from=label,values_from=c(estimate,moe) )

poverty_df_black<- poverty_df_Black %>% filter(label=='Estimate!!Total:' | label=='Estimate!!Total:!!Income in the past 12 months below poverty level:')%>%select(name, estimate,moe,label)%>%tidyr::pivot_wider(names_from=label,values_from=c(estimate,moe) )

poverty_df_black_white<- merge(poverty_df_black, poverty_df_white, by='name')

write.table(poverty_df_black_white,"clipboard", sep='\t', row.names=FALSE )

```



### Means of Transportation to Work

```{r}

white_transport_df <- get_acs_recs(geography = 'county',
             table.names = c('B08105H'),
             years=c(2019),
             acs.type = 'acs5')%>%filter(name=='Region')



asian_transport_df <- get_acs_recs(geography = 'county',
             table.names = c('B08105D'),
             years=c(2019),
             acs.type = 'acs5')%>%filter(name=='Region')



white_asian_df<-merge(asian_transport_df, white_transport_df, by='label')%>% 
  select(estimate.x, moe.x, label, estimate.y, moe.y)




write.table(white_asian_df,"clipboard", sep='\t', row.names=FALSE )


white_asian_df
```
# Asian group by county

```{r}
Asian_df_county <- get_acs_recs(geography = 'county',
             table.names = c('B02015'),
             years=c(2019),
             acs.type = 'acs5')

write.table(Asian_df_county,"clipboard", sep='\t', row.names=FALSE )
```




```{r}


```


```{r}
Asian_df_tract<- get_acs_recs(geography = 'tract',
             table.names = c('B02015'),
             years=c(2019),
             acs.type = 'acs5')

```



```{r}
library(stringr)
Asian_df_max_tract<-Asian_df_tract%>% select('GEOID', 'label', 'estimate', 'moe')%>% filter(label!='Estimate!!Total:') %>% group_by(GEOID)%>%top_n(1, estimate)%>%mutate(top_group=str_replace_all(label, 'Estimate!!Total:!!', replacement=''))%>%
             filter(estimate>=200)%>%
          mutate('GEOID'=as.character(GEOID))%>%select(GEOID, top_group, estimate)
```

```{r}
library(htmltools)
create_tract_map_categorical <- function(tbl, tract.lyr,
                             map.title = NULL, map.subtitle = NULL,
                             map.title.position = NULL,
                             legend.title = NULL, legend.subtitle = NULL,
                             map.lat=47.615, map.lon=-122.257, map.zoom=8.5, wgs84=4326){
  #print(tract.tbl)


    tract.lyr<-tract.lyr%>%
    # make geo names across 2010 and 2020
    dplyr::rename_at(dplyr::vars(matches("geoid10")),function(x) "geoid") %>%
    dplyr::rename_at(dplyr::vars(matches("geoid20")),function(x) "geoid")

  c.layer <- dplyr::left_join(tract.lyr,tbl, by = c("geoid"="GEOID")) %>%
    sf::st_transform(wgs84)

  pal <- leaflet::colorFactor(palette="Set3", domain = c.layer$top_group)

 

  labels <- paste0('Top Asian Alone Group: ', c.layer$top_group,'<br>Population: ',c.layer$estimate )%>%lapply(htmltools::HTML)

  m <- leaflet::leaflet() %>%
    leaflet::addMapPane(name = "polygons", zIndex = 410) %>%
    leaflet::addMapPane(name = "maplabels", zIndex = 500) %>% # higher zIndex rendered on top

    leaflet::addProviderTiles("CartoDB.VoyagerNoLabels") %>%
    leaflet::addProviderTiles("CartoDB.VoyagerOnlyLabels",
                              options = leaflet::leafletOptions(pane = "maplabels"),
                              group = "Labels") %>%

    leaflet::addEasyButton(leaflet::easyButton(icon="fa-globe",
                                               title="Region",
                                               onClick=leaflet::JS("function(btn, map){map.setView([47.615,-122.257],8.5); }"))) %>%
    leaflet::addPolygons(data=c.layer,
                         fillOpacity = 0.7,
                         fillColor = pal(c.layer$top_group),
                         weight = 0.7,
                         color = "#BCBEC0",
                         group="Population",
                         opacity = 0,
                         stroke=FALSE,
                         options = leaflet::leafletOptions(pane = "polygons"),
                         dashArray = "",
                         highlight = leaflet::highlightOptions(
                           weight =5,
                           color = "76787A",
                           dashArray ="",
                           fillOpacity = 0.7,
                           bringToFront = TRUE),
                         label = labels,
                         labelOptions = leaflet::labelOptions(
                           style = list("font-weight" = "normal", padding = "3px 8px"),
                           textsize = "15px",
                           direction = "auto")) %>%

    leaflet::addLegend(pal = pal,
                       values = c.layer$top_group,
                       position = "bottomright",
                       title = paste(legend.title, '<br>', legend.subtitle)) %>%

    leaflet::addControl(html = paste(map.title, '<br>', map.subtitle),
                        position = map.title.position,
                        layerId = 'mapTitle') %>%

    leaflet::addLayersControl(baseGroups = "CartoDB.VoyagerNoLabels",
                              overlayGroups = c("Labels", "Population")) %>%

    leaflet::setView(lng=map.lon, lat=map.lat, zoom=map.zoom)

  return(m)

}

```
```{r}
library(htmltools)
m<-create_tract_map_categorical(tbl=Asian_df_max_tract, tract.lyr=tract.lyr,  
                 legend.title='Top Asian Alone group', legend.subtitle=' by Census Tract')
```



# Job sector

```{r}
Asian_df <- get_acs_recs(geography = 'county',
             table.names = c('C24010D'),
             years=c(2019),
             acs.type = 'acs5')%>%filter(name=='Region')


white_df <- get_acs_recs(geography = 'county',
             table.names = c('C24010H'),
             years=c(2019),
             acs.type = 'acs5')%>%filter(name=='Region')




white_Black_df<-merge(Black_df, white_df, by='label')%>% 
  select(estimate.x, moe.x, label, estimate.y, moe.y)

write.table(white_Black_df,"clipboard", sep='\t', row.names=FALSE )
white_Black_df
```



```{r}

# to determine the threshold for reporting, I'm assuming the 5 year CS has roughly a 10% sample rate of the population, if the population of adults is 300 then the MOE is around 15%; it looks like the map ranges from shares around 10% to 97% so there is truly a wide range, and for rough story telling it should be okay.

edu<-get_acs_recs(geography = 'tract',
             table.names = 'C15002D',
             years = 2019,
             acs.type = 'acs5')%>% select('GEOID', 'label', 'estimate', 'moe')%>%filter(
label=='Estimate!!Total:!!Male:!!Bachelor\'s degree or higher' | label==
'Estimate!!Total:!!Female:!!Bachelor\'s degree or higher' | label==
'Estimate!!Total:')%>% pivot_wider(names_from=label, values_from=c(estimate, moe))%>%mutate('Total Bachelor\'s degree or higher'=`estimate_Estimate!!Total:!!Male:!!Bachelor's degree or higher` +`estimate_Estimate!!Total:!!Female:!!Bachelor's degree or higher`)%>%mutate(tots_with_not_enough_data=ifelse(`estimate_Estimate!!Total:`>300,`estimate_Estimate!!Total:`,NaN))%>% mutate('share_unformatted'= `Total Bachelor's degree or higher`/tots_with_not_enough_data)%>%mutate('estimate'=round(share_unformatted*100,0))





```
```{r}

create_tract_map_share_edu <- function(tract.tbl, tract.lyr,
                             map.title = NULL, map.subtitle = NULL,
                             map.title.position = NULL,
                             legend.title = NULL, legend.subtitle = NULL,
                             map.lat=47.615, map.lon=-122.257, map.zoom=8.5, wgs84=4326){


  # Summarize and Aggregate Tract Data by Year and Attribute to Map and join to tract layer for mapping
  # rename census value column to estimate to match ACS
  # also allow for the easy mapping of equity geographies
  tbl <- tract.tbl %>%
    dplyr::rename_at(dplyr::vars(matches("value")),function(x) "estimate") %>%
    dplyr::rename_at(dplyr::vars(matches("equity_geog_vs_50_percent")),function(x) "estimate") %>%
    dplyr::rename_at(dplyr::vars(matches('equity_geog_vs_reg_total')),function(x) "estimate") %>%
    dplyr::rename_at(dplyr::vars(matches("geoid")),function(x) "GEOID") %>%
    dplyr::select(.data$GEOID,.data$estimate) %>%
    dplyr::mutate(dplyr::across(c('GEOID'), as.character))%>%
    dplyr::group_by(.data$GEOID) %>%
    dplyr::summarise(Share=sum(.data$estimate))

    tract.lyr<-tract.lyr%>%
    # make geo names across 2010 and 2020
    dplyr::rename_at(dplyr::vars(matches("geoid10")),function(x) "geoid") %>%
    dplyr::rename_at(dplyr::vars(matches("geoid20")),function(x) "geoid")

  c.layer <- dplyr::left_join(tract.lyr,tbl, by = c("geoid"="GEOID")) %>%
    sf::st_transform(wgs84)

  pal <- leaflet::colorNumeric(palette="Purples", domain = c.layer$Share)


  labels <- paste0("Census Tract: ", c.layer$geoid, '<p></p>',
                   'Share: ', paste(c.layer$Share, "%", sep=""))%>% lapply(htmltools::HTML)

  m <- leaflet::leaflet() %>%
    leaflet::addMapPane(name = "polygons", zIndex = 410) %>%
    leaflet::addMapPane(name = "maplabels", zIndex = 500) %>% # higher zIndex rendered on top

    leaflet::addProviderTiles("CartoDB.VoyagerNoLabels") %>%
    leaflet::addProviderTiles("CartoDB.VoyagerOnlyLabels",
                              options = leaflet::leafletOptions(pane = "maplabels"),
                              group = "Labels") %>%

    leaflet::addEasyButton(leaflet::easyButton(icon="fa-globe",
                                               title="Region",
                                               onClick=leaflet::JS("function(btn, map){map.setView([47.615,-122.257],8.5); }"))) %>%
    leaflet::addPolygons(data=c.layer,
                         fillOpacity = 0.7,
                         fillColor = pal(c.layer$Share),
                         weight = 0.7,
                         color = "#BCBEC0",
                         group="Population",
                         opacity = 0,
                         stroke=FALSE,
                         options = leaflet::leafletOptions(pane = "polygons"),
                         dashArray = "",
                         highlight = leaflet::highlightOptions(
                           weight =5,
                           color = "76787A",
                           dashArray ="",
                           fillOpacity = 0.7,
                           bringToFront = TRUE),
                         label = labels,
                         labelOptions = leaflet::labelOptions(
                           style = list("font-weight" = "normal", padding = "3px 8px"),
                           textsize = "15px",
                           direction = "auto")) %>%
    leaflet::addLegend(pal = pal,
                       values = c.layer$Share,
                       position = "bottomright",
                       title = paste(legend.title, '<br>', legend.subtitle),
                       labFormat=leaflet::labelFormat(suffix='%'),
                       na.label='not enough <br>     data to report')%>%

    leaflet::addControl(html = paste(map.title, '<br>', map.subtitle),
                        position = map.title.position,
                        layerId = 'mapTitle') %>%

    leaflet::addLayersControl(baseGroups = "CartoDB.VoyagerNoLabels",
                              overlayGroups = c("Labels", "Population")) %>%

    leaflet::setView(lng=map.lon, lat=map.lat, zoom=map.zoom)

  return(m)

}
```

```{r}

m2<-create_tract_map_share_edu(tract.tbl=edu, tract.lyr=tract.lyr,  
                 legend.title='Share of Asian Alone Adults', legend.subtitle=' with Bachelor\`s degree or higher')
```




```{r}
edu_df_tract <- get_acs_recs(geography = 'tract',
             table.names = c('B15002'),
             years=c(2019),
             acs.type = 'acs5')


```




```{r}
inc_race_srvyr_obj<-get_psrc_pums(span=5, dyear=2019, level="p", vars=c("PERNP","RAC2P", "WKHP"))%>% filter(WKHP>10)

inc_race<-psrc_pums_median(inc_race_srvyr_obj, stat_var="PERNP",group_vars = c("RAC2P"))
#%>%filter(COUNTY=='Region')%>%pivot_wider(names_from= RAC2P, values_from=c(PERNP_median, PERNP_median_moe))

write.table(inc_race,"clipboard-16384", sep='\t', row.names=FALSE )

```


```{r}
inc_race_srvyr_obj<-get_psrc_pums(span=5, dyear=2019, level="p", vars=c("PERNP","RAC1P", "WKHP"))%>% filter(WKHP>10)

inc_race<-psrc_pums_median(inc_race_srvyr_obj, stat_var="PERNP",group_vars = c("RAC1P"))
#%>%filter(COUNTY=='Region')%>%pivot_wider(names_from= RAC2P, values_from=c(PERNP_median, PERNP_median_moe))

write.table(inc_race,"clipboard-16384", sep='\t', row.names=FALSE )

```


```{r}
employment_status_sex_srvyr_obj<-get_psrc_pums(span=5, dyear=2019, level="p", vars=c("SEX","ESR", 'WKHP'))



emp_sex<-psrc_pums_count(employment_status_sex_srvyr_obj,group_vars = c("SEX", "ESR"))%>%filter(COUNTY=='Region')

write.table(emp_sex,"clipboard-16384", sep='\t', row.names=FALSE )

```


```{r}
jobs_df <- get_acs_recs(geography = 'county',
             table.names = c('C24030'),
             years=c(2019),
             acs.type = 'acs5')%>%filter(name=='Region')%>% arrange(label, name)%>%select(name, label, estimate, moe)
jobs_df
write.table(jobs_df,"clipboard-16384", sep='\t', row.names=FALSE )
```

```{r}
inc_sex_srvyr_obj<-get_psrc_pums(span=5, dyear=2019, level="p", vars=c("SEX","PERNP","WKHP", "INDP"))%>% filter(WKHP>35)%>% mutate(short_industry=str_extract(INDP,"[aA-zZ]{1,4}" ))

inc_sex<-psrc_pums_median(inc_sex_srvyr_obj, target_var='PERNP',group_vars = c("SEX", "short_industry"))%>%filter(COUNTY=='Region')%>%pivot_wider(names_from= c(SEX), values_from=c(PERNP_median, PERNP_median_moe))
write.table(inc_sex,"clipboard-16384", sep='\t', row.names=FALSE )
```
```