---
title: "Womens History Month ACS/PUMS Data"
author: "suzanne"
date: "3/2022"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Notes for Asian American Pacific Islander month

```{r load_libraries}
library(devtools)
library(sf)
library(dplyr)
library(psrccensus)
library(ggplot2)
library(tidycensus)
library(tidyr)
Sys.getenv("CENSUS_API_KEY")
```
```{r}
acs_5_vars<-load_variables(2020, 'acs5', cache=TRUE)

```


	
B01002D_001 Median Age
	
B02015    specific origins

B03002 Hispanic
	
B08105 Mode to Work
	
B10051 Grandparents
	
B17001 Poverty
	
C15002 Education
	
B16005 English proficiency

RAC1P
RAC2P
RAC3P detailed race in combination
RACASN- alone or in combination


HINCP

HUGCL
HUPAC


### Educational Attainment
compare 2020 and 2019 5 year
```{r}




Asian_df <- get_acs_recs(geography = 'county',
             table.names = c('C15002D'),
             years=c(2020),
             acs.type = 'acs5')%>%filter(name=='Region')

PI_df <- get_acs_recs(geography = 'county',
             table.names = c('C15002E'),
             years=c(2020),
             acs.type = 'acs5')%>%filter(name=='Region')

white_df <- get_acs_recs(geography = 'county',
             table.names = c('C15002H'),
             years=c(2020),
             acs.type = 'acs5')%>%filter(name=='Region')


Asian_PI_df<-merge(Asian_df, PI_df, by='label')
Asian_PI_white_df<-merge(Asian_PI_df, white_df, by='label')


Asian_PI_white_df_formatted<-Asian_PI_white_df %>% select(label,estimate.x, moe.x, estimate.y, moe.y, estimate.y, moe.y, estimate, moe) %>% rename(Asian_Educational_Attainment=estimate.x, Asian_Educational_Attainment_MOE=moe.x, Pacific_Islander_Educational_Attainment=estimate.y, Pacific_Islander_Educational_Attainment_MOE=moe.y, White_Educational_Attainment=estimate, White_Educational_Attainment_MOE=moe)





write.table(Asian_PI_white_df_formatted,"clipboard", sep='\t', row.names=FALSE )


Asian_PI_white_df_formatted


```

```{r}




Asian_df <- get_acs_recs(geography = 'county',
             table.names = c('C15002D'),
             years=c(2019),
             acs.type = 'acs5')%>%filter(name=='Region')

PI_df <- get_acs_recs(geography = 'county',
             table.names = c('C15002E'),
             years=c(2019),
             acs.type = 'acs5')%>%filter(name=='Region')

white_df <- get_acs_recs(geography = 'county',
             table.names = c('C15002H'),
             years=c(2019),
             acs.type = 'acs5')%>%filter(name=='Region')


Asian_PI_df<-merge(Asian_df, PI_df, by='label')
Asian_PI_white_df<-merge(Asian_PI_df, white_df, by='label')


Asian_PI_white_df_formatted<-Asian_PI_white_df %>% select(label,estimate.x, moe.x, estimate.y, moe.y, estimate.y, moe.y, estimate, moe) %>% rename(Asian_Educational_Attainment=estimate.x, Asian_Educational_Attainment_MOE=moe.x, Pacific_Islander_Educational_Attainment=estimate.y, Pacific_Islander_Educational_Attainment_MOE=moe.y, White_Educational_Attainment=estimate, White_Educational_Attainment_MOE=moe)





write.table(Asian_PI_white_df_formatted,"clipboard", sep='\t', row.names=FALSE )


Asian_PI_white_df_formatted


```


by county
```{r}




Asian_df <- get_acs_recs(geography = 'county',
             table.names = c('C15002D'),
             years=c(2020),
             acs.type = 'acs5')

PI_df <- get_acs_recs(geography = 'county',
             table.names = c('C15002E'),
             years=c(2020),
             acs.type = 'acs5')

white_df <- get_acs_recs(geography = 'county',
             table.names = c('C15002H'),
             years=c(2020),
             acs.type = 'acs5')


Asian_PI_df<-merge(Asian_df, PI_df, by=c('label', 'name'))
Asian_PI_white_df<-merge(Asian_PI_df, white_df, by=c('label', 'name'))


Asian_PI_white_df_formatted<-Asian_PI_white_df %>% select(name,label,estimate.x, moe.x, estimate.y, moe.y, estimate.y, moe.y, estimate, moe) %>% rename(Asian_Educational_Attainment=estimate.x, Asian_Educational_Attainment_MOE=moe.x, Pacific_Islander_Educational_Attainment=estimate.y, Pacific_Islander_Educational_Attainment_MOE=moe.y, White_Educational_Attainment=estimate, White_Educational_Attainment_MOE=moe)





write.table(Asian_PI_white_df_formatted,"clipboard", sep='\t', row.names=FALSE )


Asian_PI_white_df_formatted


### Where were people born?
```
2020
```{r}


nativity_df_Asian<- get_acs_recs(geography = 'county',
             table.names = c('B06004D'),
             years=c(2020),
             acs.type = 'acs5')



nativity_df_Asian_formatted<-nativity_df_Asian 

nativity_df_Asian_formatted
write.table(nativity_df_Asian_formatted,"clipboard", sep='\t', row.names=FALSE )

```
2019
```{r}


nativity_df_Asian<- get_acs_recs(geography = 'county',
             table.names = c('B06004D'),
             years=c(2019),
             acs.type = 'acs5')



nativity_df_Asian_formatted<-nativity_df_Asian 

nativity_df_Asian_formatted
write.table(nativity_df_Asian_formatted,"clipboard", sep='\t', row.names=FALSE )

```
```{r}
nativity_df_PI<- get_acs_recs(geography = 'county',
             table.names = c('B06004E'),
             years=c(2020),
             acs.type = 'acs5')



nativity_df_PI_formatted<-nativity_df_PI

nativity_df_PI_formatted
write.table(nativity_df_PI_formatted,"clipboard", sep='\t', row.names=FALSE )


```
```{r}
nativity_df_PI<- get_acs_recs(geography = 'county',
             table.names = c('B06004E'),
             years=c(2019),
             acs.type = 'acs5')



nativity_df_PI_formatted<-nativity_df_PI

nativity_df_PI_formatted
write.table(nativity_df_PI_formatted,"clipboard", sep='\t', row.names=FALSE )


```

### Median Income
B19013D
```{r}




Asian_df <- get_acs_recs(geography = 'county',
             table.names = c('B19013D'),
             years=c(2020),
             acs.type = 'acs5')
PI_df <- get_acs_recs(geography = 'county',
             table.names = c('B19013E'),
             years=c(2020),
             acs.type = 'acs5')

white_df <- get_acs_recs(geography = 'county',
             table.names = c('B19013H'),
             years=c(2020),
             acs.type = 'acs5')


Asian_PI_df<-merge(Asian_df, PI_df, by=c('label', 'name'))
Asian_PI_white_df<-merge(Asian_PI_df, white_df, by=c('label', 'name'))


Asian_PI_white_df_formatted<-Asian_PI_white_df %>% select(label,name, estimate.x, moe.x, estimate.y, moe.y, estimate.y, moe.y, estimate, moe) %>% rename(Asian_Educational_Attainment=estimate.x, Asian_Median_Income_MOE=moe.x, Pacific_Islander_Median_Income=estimate.y, Pacific_Islander_Median_Income_MOE=moe.y, White_Median_Income=estimate, White_Median_Income_MOE=moe)





write.table(Asian_PI_white_df_formatted,"clipboard", sep='\t', row.names=FALSE )


Asian_PI_white_df_formatted
```
### Median Income
B19013D
2019
```{r}




Asian_df <- get_acs_recs(geography = 'county',
             table.names = c('B19013D'),
             years=c(2019),
             acs.type = 'acs5')
PI_df <- get_acs_recs(geography = 'county',
             table.names = c('B19013E'),
             years=c(2019),
             acs.type = 'acs5')

white_df <- get_acs_recs(geography = 'county',
             table.names = c('B19013H'),
             years=c(2019),
             acs.type = 'acs5')


Asian_PI_df<-merge(Asian_df, PI_df, by=c('label', 'name'))
Asian_PI_white_df<-merge(Asian_PI_df, white_df, by=c('label', 'name'))


Asian_PI_white_df_formatted<-Asian_PI_white_df %>% select(label,name, estimate.x, moe.x, estimate.y, moe.y, estimate.y, moe.y, estimate, moe) %>% rename(Asian_Educational_Attainment=estimate.x, Asian_Median_Income_MOE=moe.x, Pacific_Islander_Median_Income=estimate.y, Pacific_Islander_Median_Income_MOE=moe.y, White_Median_Income=estimate, White_Median_Income_MOE=moe)





write.table(Asian_PI_white_df_formatted,"clipboard", sep='\t', row.names=FALSE )


Asian_PI_white_df_formatted
```
## Homeownership

```{r}


homeownership_black<-get_acs_recs(geography = 'county',
             table.names = c('B25003B'),
             years=c(2019),
             acs.type = 'acs1')

homeownership_white<-get_acs_recs(geography = 'county',
             table.names = c('B25003H'),
             years=c(2019),
             acs.type = 'acs1')


homeownership_df_black_white<- merge(homeownership_black, homeownership_white, by=c('label','name'))
write.table(homeownership_df_black_white,"clipboard", sep='\t', row.names=FALSE )

homeownership_df_black_white

```

### Median Age by Black and All

```{r}

Black_age_df<-get_acs_recs(geography = 'county',
             table.names = c('B01002B'),
             years=c(2019),
             acs.type = 'acs5')


white_age_df <- get_acs_recs(geography = 'county',
             table.names = c('B01002H'),
             years=c(2019),
             acs.type = 'acs5')


white_Black_df<-merge(Black_age_df, white_age_df, by=c('name','label'))%>% filter(label=='Estimate!!Median age --!!Total:')%>%
  select(name,estimate.x, estimate.y, )%>% rename('County'='name', 'Black population' ='estimate.x', 'White Population'='estimate.y')

write.table(white_Black_df,"clipboard", sep='\t', row.names=FALSE )
white_Black_df

```



### Poverty Rates

```{r}


poverty_df_Black<- get_acs_recs(geography = 'county',
             table.names = c('B17020B'),
             years=c(2019),
             acs.type = 'acs5')
poverty_df_white <- get_acs_recs(geography = 'county',
             table.names = c('B17020H'),
             years=c(2019),
             acs.type = 'acs5')

poverty_df_white<- poverty_df_white %>% filter(label=='Estimate!!Total:' | label=='Estimate!!Total:!!Income in the past 12 months below poverty level:')%>%select(name, estimate,moe,label)%>%tidyr::pivot_wider(names_from=label,values_from=c(estimate,moe) )

poverty_df_black<- poverty_df_Black %>% filter(label=='Estimate!!Total:' | label=='Estimate!!Total:!!Income in the past 12 months below poverty level:')%>%select(name, estimate,moe,label)%>%tidyr::pivot_wider(names_from=label,values_from=c(estimate,moe) )

poverty_df_black_white<- merge(poverty_df_black, poverty_df_white, by='name')

write.table(poverty_df_black_white,"clipboard", sep='\t', row.names=FALSE )

```



### Means of Transportation to Work

```{r}

white_transport_df <- get_acs_recs(geography = 'county',
             table.names = c('B08105H'),
             years=c(2020),
             acs.type = 'acs5')%>%filter(name=='Region')



asian_transport_df <- get_acs_recs(geography = 'county',
             table.names = c('B08105D'),
             years=c(2020),
             acs.type = 'acs5')%>%filter(name=='Region')



white_asian_df<-merge(asian_transport_df, white_transport_df, by='label')%>% 
  select(estimate.x, moe.x, label, estimate.y, moe.y)




write.table(white_asian_df,"clipboard", sep='\t', row.names=FALSE )


white_asian_df
```

### Means of Transportation to Work

```{r}

white_transport_df <- get_acs_recs(geography = 'county',
             table.names = c('B08105A'),
             years=c(2020),
             acs.type = 'acs5')%>%filter(name=='Region')



pi_transport_df <- get_acs_recs(geography = 'county',
             table.names = c('B08105E'),
             years=c(2020),
             acs.type = 'acs5')%>%filter(name=='Region')



white_pi_df<-merge(pi_transport_df, white_transport_df, by='label')%>% 
  select(estimate.x, moe.x, label, estimate.y, moe.y)




write.table(white_pi_df,"clipboard", sep='\t', row.names=FALSE )

white_pi_df
```


### B25014, Occupants per room
```{r}
white_df <- get_acs_recs(geography = 'county',
             table.names = c('B25014H'),
             years=c(2019),
             acs.type = 'acs5')



Black_df <- get_acs_recs(geography = 'county',
             table.names = c('B25014B'),
             years=c(2019),
             acs.type = 'acs5')
```




```{r}
white_Black_df<-merge(Black_df, white_df, by='label')%>% 
  select(estimate.x, moe.x, label, estimate.y, moe.y)




write.table(white_Black_df,"clipboard", sep='\t', row.names=FALSE )


white_Black_df
```

```{r}
white_df <- get_acs_recs(geography = 'county',
             table.names = c('B25032H'),
             years=c(2019),
             acs.type = 'acs5')%>%filter(name=='Region')



Black_df <- get_acs_recs(geography = 'county',
             table.names = c('B25032B'),
             years=c(2019),
             acs.type = 'acs5')%>%filter(name=='Region')

white_Black_df<-merge(Black_df, white_df, by='label')%>% 
  select(estimate.x, moe.x, label, estimate.y, moe.y)

write.table(white_Black_df,"clipboard", sep='\t', row.names=FALSE )%>%filter(name=='Region')
```

# Computers at home
```{r}
Black_df <- get_acs_recs(geography = 'county',
             table.names = c('B28009B'),
             years=c(2019),
             acs.type = 'acs5')%>%filter(name=='Region')


white_df <- get_acs_recs(geography = 'county',
             table.names = c('B28009H'),
             years=c(2019),
             acs.type = 'acs5')%>%filter(name=='Region')




white_Black_df<-merge(Black_df, white_df, by='label')%>% 
  select(estimate.x, moe.x, label, estimate.y, moe.y)

write.table(white_Black_df,"clipboard", sep='\t', row.names=FALSE )%>%filter(name=='Region')
```
# Job sector

```{r}
Black_df <- get_acs_recs(geography = 'county',
             table.names = c('C24010B'),
             years=c(2019),
             acs.type = 'acs5')%>%filter(name=='Region')


white_df <- get_acs_recs(geography = 'county',
             table.names = c('C24010H'),
             years=c(2019),
             acs.type = 'acs5')%>%filter(name=='Region')




white_Black_df<-merge(Black_df, white_df, by='label')%>% 
  select(estimate.x, moe.x, label, estimate.y, moe.y)

write.table(white_Black_df,"clipboard", sep='\t', row.names=FALSE )
white_Black_df
```


# Health Insurance
```{r}
Black_df <- get_acs_recs(geography = 'county',
             table.names = c('C27001B'),
             years=c(2019),
             acs.type = 'acs5')%>%filter(name=='Region')


white_df <- get_acs_recs(geography = 'county',
             table.names = c('C27001H'),
             years=c(2019),
             acs.type = 'acs5')%>%filter(name=='Region')




white_Black_df<-merge(Black_df, white_df, by='label')%>% 
  select(estimate.x, moe.x, label, estimate.y, moe.y)
white_Black_df

write.table(white_Black_df,"clipboard", sep='\t', row.names=FALSE )
```

```{r}
foreign_born<-get_acs_recs(geography = 'tract',
             table.names = 'B05006',
             years = 2019,
             acs.type = 'acs5')

foreign_born_Africa <- foreign_born %>% filter(label=='Estimate!!Total:!!Africa:')

             
gdb.nm <- paste0("MSSQL:server=",
"AWS-PROD-SQL\\Sockeye",
";database=",
"ElmerGeo",
";trusted_connection=yes")

spn <-  2285
wgs84=4326

tract_layer_name <- "dbo.tract2010_nowater"

tract.lyr <- st_read(gdb.nm, tract_layer_name, crs = spn)



m<-psrccensus::create_tract_map(tract.tbl=foreign_born_Africa, tract.lyr=tract.lyr,  
                 legend.title='African Foreign Born Population', legend.subtitle='by Census Tract')




Map of difference in median income men vs women

### Median Earnings Tract
```{r}
earnings_df_tract <- get_acs_recs(geography = 'tract',
             table.names = c('B20017'),
             years=c(2019),
             acs.type = 'acs5')%>%filter(label=="Estimate!!Median earnings in the past 12 months (in 2019 inflation-adjusted dollars) --!!Total (dollars):!!Male --!!Total (dollars)!!Worked full-time, year-round in the past 12 months (dollars)" | label=='Estimate!!Median earnings in the past 12 months (in 2019 inflation-adjusted dollars) --!!Total (dollars):!!Female --!!Total (dollars)!!Worked full-time, year-round in the past 12 months (dollars)')


```
``


###Educational attainment
```{r}
edu_df <- get_acs_recs(geography = 'county',
             table.names = c('B15002'),
             years=c(2019),
             acs.type = 'acs5')%>%filter(name=='Region')%>% arrange(label, name)%>%select(name, label, estimate, moe)
edu_df
write.table(edu_df,"clipboard-16384", sep='\t', row.names=FALSE )
```