---
title: "Black History Month ACS Data"
author: "suzanne"
date: "8/19/2021"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Notes for Black History Month

```{r load_libraries}
library(devtools)
library(sf)
library(dplyr)
library(psrccensus)
library(ggplot2)
library(tidycensus)
Sys.getenv("CENSUS_API_KEY")
```

I searched the api table list for AIAN to see what data is available.
<https://api.census.gov/data/2019/acs/acs5/variables.html>

How does AIAN age breakdown compare to non-AIAN? \* Age is found in
B1002I, B1002A

What Races do AIAN people identify with? \* B3002

What specific origins do people come from? \* B300, B05003

What means of transportation do people use to get to work? \* B06004,
B06007 means of transportation to work for workplace geo

Information about grandparents living with grandchildren \* B08105H; I
grandparents B10051I

Are more AIAN women giving birth than all women? \* B13002 women who
gave birth

What are the educational attainment levels for AIAN people? \* B15002
educational attainment

What are the poverty rates for AIAN people? \* B17001 B19001

```{r}
var_list<- load_variables(2019, "acs5", cache = TRUE)
```


```{r}
homeownership<-get_acs_recs(geography = 'county',
             table.names = c('B25003'),
             years=c(2019),
             acs.type = 'acs1')
write.table(homeownership,"clipboard", sep='\t', row.names=FALSE )


homeownership_white<-get_acs_recs(geography = 'county',
             table.names = c('B25003H'),
             years=c(2019),
             acs.type = 'acs1')
write.table(homeownership_white,"clipboard", sep='\t', row.names=FALSE )

homeownership_black<-get_acs_recs(geography = 'county',
             table.names = c('B25003C'),
             years=c(2019),
             acs.type = 'acs1')
write.table(homeownership_black, "clipboard", sep='\t', row.names=FALSE )
```

### Median Age by AIAN and All

```{r}

overall_age_df <- get_acs_recs(geography = 'county',
             table.names = c('B01002'),
             years=c(2019),
             acs.type = 'acs5')

AIAN_age_df<-get_acs_recs(geography = 'county',
             table.names = c('B01002C'),
             years=c(2019),
             acs.type = 'acs5')

AIAN_age_df_test <- AIAN_age_df%>% filter(name !='Region' | (!(stringr::str_detect(label, 'Median'))&(stringr::str_detect(label, 'Average'))))

AIAN_age_df_test2 <- AIAN_age_df%>% filter(!(stringr::str_detect(label, 'Median')))
                                     
                                     #||
                             #!(stringr::str_detect(label, 'Median')& (stringr::str_detect(label,'Average'))))

overall_AIAN_df<-rbind(overall_age_df, AIAN_age_df)
age_df<-overall_AIAN_df%>% filter(label=='Estimate!!Median age --!!Total:' & name !='Region') %>% mutate('AIAN'=ifelse(variable=='B01002_001', 'All Population', 'AIAN Population'))
write.table(age_df,"clipboard", sep='\t', row.names=FALSE )

age_df

ggplot(data=age_df, aes(x=name, y=estimate, fill=AIAN))+geom_bar(stat='identity', position=position_dodge())+
  geom_text(aes(label=estimate), vjust=1.6, color="white",
            position = position_dodge(0.9), size=3.5)

```

### Women who gave birth

```{r}
women_birth_df<-get_acs_recs(geography = 'county',
             table.names = c('B13002'),
             years=c(2019),
             acs.type = 'acs1') 
women_birth_df_AIAN<-get_acs_recs(geography = 'county',
             table.names = c('B13002C'),
             years=c(2019),
             acs.type = 'acs5')
women_birth_df
women_birth_df_AIAN
write.table(women_birth_df,"clipboard", sep='\t', row.names=FALSE )
write.table(women_birth_df_AIAN,"clipboard", sep='\t', row.names=FALSE )
```

### Poverty Rates

```{r}
poverty_df_white <- get_acs_recs(geography = 'county',
             table.names = c('B17020H'),
             years=c(2019),
             acs.type = 'acs5')


poverty_df_AIAN<- get_acs_recs(geography = 'county',
             table.names = c('B17020C'),
             years=c(2019),
             acs.type = 'acs5')

write.table(poverty_df_white,"clipboard", sep='\t', row.names=FALSE )
write.table(poverty_df_AIAN,"clipboard", sep='\t', row.names=FALSE )
poverty_df_white
poverty_df_AIAN
```

17% of AIAN people in poverty, vs 7% of whites

### Median Income

```{r}

income_df_white <- get_acs_recs(geography = 'county',
             table.names = c('B19013H'),
             years=c(2019),
             acs.type = 'acs5')


income_df_AIAN<- get_acs_recs(geography = 'county',
             table.names = c('B19013C'),
             years=c(2019),
             acs.type = 'acs5')

white_AIAN_df<-merge(income_df_white, income_df_AIAN, by ='name')

white_AIAN_df
write.table(white_AIAN_df,"clipboard", sep='\t', row.names=FALSE )

```

### Means of Transportation to Work

```{r}

white_transport_df <- get_acs_recs(geography = 'county',
             table.names = c('B08105A'),
             years=c(2019),
             acs.type = 'acs5')



AIAN_transport_df <- get_acs_recs(geography = 'county',
             table.names = c('B08105C'),
             years=c(2019),
             acs.type = 'acs5')



white_AIAN_df<-rbind(white_transport_df, AIAN_transport_df)%>%filter(name=='Region')



write.table(white_AIAN_df,"clipboard", sep='\t', row.names=FALSE )


white_AIAN_df



```

Which tribes?

```{r}
AIAN_tribes<- get_acs_recs(geography = 'county',
             table.names = c('B02014'),
             years=c(2019),
             acs.type = 'acs5')
write.table(AIAN_tribes,"clipboard", sep='\t', row.names=FALSE )
```

34,176 total people who identify as American Indian or Alaskan Native

24,600 identified American Indian Tribes

3,239 Alaskan Native Tribes

```{r}
race_table<-get_acs_recs(geography = 'county',
             table.names = c('B02001'),
             years=c(2019),
             acs.type = 'acs1')

race_format<-format_estimates(race_table, type = 'share', moe = TRUE)
```



