---
title: "hispanic_heritage"
author: "suzanne"
date: "8/19/2021"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Notes

These are notes for Hispanic heritage month.
<https://www.hispanicheritagemonth.gov/> September 15 to October 15

```{r load_libraries}
library(devtools)
library(sf)
library(dplyr)
library(psrccensus)
library(ggplot2)
Sys.getenv("CENSUS_API_KEY")
```

How does Hispanic age breakdown compare to non-Hispanic?

age B1002I, B1002A what Races do Hispanic people identify as B3002; use
Census what specific origins do people come from B3001 nativity B05003;
B06004, B06007 means of transportation to work for workplace geo
B08105H; I grandparents B10051I

B11001I B13002 women who gave birth B15002 educational attainment B17001
poverty B19001 B24010 occupation

```{r}
overall_age_df <- get_acs_recs(geography = 'county',
             table.names = c('B01002'),
             years=c(2019),
             acs.type = 'acs1')

hispanic_age_df<-get_acs_recs(geography = 'county',
             table.names = c('B01002I'),
             years=c(2019),
             acs.type = 'acs1')

overall_hispanic_df<-rbind(overall_age_df, hispanic_age_df)
age_df<-overall_hispanic_df%>% filter(label=='Estimate!!Median age --!!Total:' & name !='Region') %>% mutate('Hispanic'=ifelse(variable=='B01002_001', 'All Population', 'Hispanic or Latino Population'))


ggplot(data=age_df, aes(x=name, y=estimate, fill=Hispanic))+geom_bar(stat='identity', position=position_dodge())+
  geom_text(aes(label=estimate), vjust=1.6, color="white",
            position = position_dodge(0.9), size=3.5)
```

Map hispanic population by tract

```{r}
tract.big.tbl <- psrccensus::get_decennial_recs(geography='tract',table_codes=c('P005'),year=c(2010))



tract.tbl <- tract.big.tbl %>%
filter(label=='Total!!Hispanic or Latino')

gdb.nm <- paste0("MSSQL:server=",
"AWS-PROD-SQL\\Sockeye",
";database=",
"ElmerGeo",
";trusted_connection=yes")

spn <-  2285

tract_layer_name <- "dbo.tract2010_nowater"

tract.lyr <- st_read(gdb.nm, tract_layer_name, crs = spn)
m<-create_tract_map(tract.tbl=tract.tbl, tract.lyr=tract.lyr,  
                 legend.title='Hispanic Population', legend.subtitle='by Census Tract')

```

```{r}
women_birth_df<-get_acs_recs(geography = 'county',
             table.names = c('B13002'),
             years=c(2019),
             acs.type = 'acs1') 
women_birth_df_hispanic<-get_acs_recs(geography = 'county',
             table.names = c('B13002I'),
             years=c(2019),
             acs.type = 'acs1')
```

1049422 women age 15 to 50 in the region

55604 gave birth (5.3%)

118645 Hispanic women

6694 (5.6%)

```{r}
country_df<-get_acs_recs(geography = 'county',
             table.names = c('B03001'),
             years=c(2019),
             acs.type = 'acs1') 
write.table(country_df, "clipboard", sep="\t", row.names=FALSE)

#ggplot(data=country_df, aes(x=label, y=estimate))+geom_bar(stat#='identity', position=position_dodge())+
#  geom_text(aes(label=estimate), vjust=1.6, color="white",
#            position = position_dodge(0.9), size=3.5)

```

Comparing 2010 and 2020 Census

Totals by County

Tract Map; Tract Difference Map
