---
title: "Hispanic-Latinx Heritage Month ACS Data"
author: "suzanne"
date: "8/19/2021"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Notes for Hispanic-Latinx heritage month

These are notes for Hispanic heritage month.
<https://www.hispanicheritagemonth.gov/> September 15 to October 15

```{r load_libraries}
library(devtools)
library(sf)
library(dplyr)
library(psrccensus)
library(ggplot2)
Sys.getenv("CENSUS_API_KEY")
```

I searched the api table list for Hispanic to see what data is
available. <https://api.census.gov/data/2019/acs/acs5/variables.html>

How does Hispanic age breakdown compare to non-Hispanic? \* Age is found
in B1002I, B1002A

What Races do Hispanic people identify with? \* B3002

What specific origins do people come from? \* B300, B05003

What means of transportation do people use to get to work? \* B06004,
B06007 means of transportation to work for workplace geo

Information about grandparents living with grandchildren \* B08105H; I
grandparents B10051I

Are more Hispanic women giving birth than all women? \* B13002 women who
gave birth

What are the educational attainment levels for Hispanic people? \*
B15002 educational attainment

What are the poverty rates for Hispanic people? \* B17001 B19001

### Median Age by Hispanic and All

```{r}

overall_age_df <- get_acs_recs(geography = 'county',
             table.names = c('B01002'),
             years=c(2019),
             acs.type = 'acs1')

hispanic_age_df<-get_acs_recs(geography = 'county',
             table.names = c('B01002I'),
             years=c(2019),
             acs.type = 'acs1')

overall_hispanic_df<-rbind(overall_age_df, hispanic_age_df)
age_df<-overall_hispanic_df%>% filter(label=='Estimate!!Median age --!!Total:' & name !='Region') %>% mutate('Hispanic'=ifelse(variable=='B01002_001', 'All Population', 'Hispanic or Latino Population'))
write.table(age_df,"clipboard", sep='\t', row.names=FALSE )

age_df

ggplot(data=age_df, aes(x=name, y=estimate, fill=Hispanic))+geom_bar(stat='identity', position=position_dodge())+
  geom_text(aes(label=estimate), vjust=1.6, color="white",
            position = position_dodge(0.9), size=3.5)

```

### Hispanic Population by Tract

```{r}
tract.big.tbl <- psrccensus::get_decennial_recs(geography='tract',table_codes=c('P005'),year=c(2010))



tract.tbl <- tract.big.tbl %>%
filter(label=='Total!!Hispanic or Latino')

gdb.nm <- paste0("MSSQL:server=",
"AWS-PROD-SQL\\Sockeye",
";database=",
"ElmerGeo",
";trusted_connection=yes")

spn <-  2285

tract_layer_name <- "dbo.tract2010_nowater"

tract.lyr <- st_read(gdb.nm, tract_layer_name, crs = spn)
m<-create_tract_map(tract.tbl=tract.tbl, tract.lyr=tract.lyr,  
                 legend.title='Hispanic Population', legend.subtitle='by Census Tract')

m
```

### Women who gave birth

```{r}
women_birth_df<-get_acs_recs(geography = 'county',
             table.names = c('B13002'),
             years=c(2019),
             acs.type = 'acs1') 
women_birth_df_hispanic<-get_acs_recs(geography = 'county',
             table.names = c('B13002I'),
             years=c(2019),
             acs.type = 'acs1')
women_birth_df
women_birth_df_hispanic
```

1049422 women age 15 to 50 in the region

55604 gave birth (5.3%)

118645 Hispanic women

6694 (5.6%)

### Poverty Rates

```{r}
poverty_df_white <- get_acs_recs(geography = 'county',
             table.names = c('B17020H'),
             years=c(2019),
             acs.type = 'acs5')


poverty_df_hispanic<- get_acs_recs(geography = 'county',
             table.names = c('B17020I'),
             years=c(2019),
             acs.type = 'acs5')


poverty_df_white
poverty_df_hispanic
```

### Median Income

```{r}

income_df_white <- get_acs_recs(geography = 'county',
             table.names = c('B19013H'),
             years=c(2019),
             acs.type = 'acs1')


income_df_hispanic<- get_acs_recs(geography = 'county',
             table.names = c('B19013I'),
             years=c(2019),
             acs.type = 'acs1')

white_hispanic_df<-merge(income_df_white, income_df_hispanic, by ='name')

white_hispanic_df
write.table(white_hispanic_df,"clipboard", sep='\t', row.names=FALSE )

```

### Means of Transportation to Work

```{r}

white_transport_df <- get_acs_recs(geography = 'county',
             table.names = c('B08105A'),
             years=c(2019),
             acs.type = 'acs5')



hispanic_transport_df <- get_acs_recs(geography = 'county',
             table.names = c('B08105I'),
             years=c(2019),
             acs.type = 'acs5')



white_hispanic_df<-rbind(white_transport_df, hispanic_transport_df)%>%filter(name=='Region')



write.table(white_hispanic_df,"clipboard", sep='\t', row.names=FALSE )


white_hispanic_df



```

### Country of Origin

```{r}

country_df_hispanic<- get_acs_recs(geography = 'county',
             table.names = c('B03001'),
             years=c(2019),
             acs.type = 'acs1')






country_df_hispanic

```
